---
title: "Data Exploration: IUCN and AquaMaps"
author: "Jamie Afflerbach"
date: "9/2/2015"
output: html_document
---

```{r,echo=F,warning=FALSE,message=F}
library(ggplot2)
library(maps)
library(readr)
library(raster)
library(tidyr)
library(dplyr)

dir_N <- c('Windows' = '//neptune.nceas.ucsb.edu/data_edit',
           'Darwin'  = '/Volumes/data_edit',
           'Linux'   = '/var/data/ohi')[[ Sys.info()[['sysname']] ]]

dir_anx      <- file.path(dir_N, 'git-annex/globalprep/SPP_ICO')
```

```{r,echo=F,warning=FALSE,message=F,results='hide'}
iucn_spp_cells <- read_csv(file.path(dir_anx, 'explore_am_v_iucn/iucn_spp_cells.csv'), col_types = 'idiii', progress = TRUE)

am_spp_cells   <- read_csv(file.path(dir_anx, 'explore_am_v_iucn/am_spp_cells.csv'), progress = TRUE)

#species list
spp_list <- read.csv('shiny/data/spp_am_v_iucn.csv', stringsAsFactors = FALSE)


#data for raster (used to plot species)

loiczid_raster_file  <- 'shiny/data/loiczid_raster.grd'
loiczid_raster       <- raster(loiczid_raster_file)
names(loiczid_raster) <- 'loiczid'
```

```{r,echo=F,warning=FALSE,message=F}

### Species Map Function ###
# This function takes a single species scientific name as input, then grabs all 
# occurrence cells and associated Aquamaps probability and/or IUCN proportional area
# per cell
get_spp_map <- function(species){
  spp_id <- spp_list %>%
    filter(sciname == species & is.na(parent_sid)) %>%
    dplyr::select(am_sid, iucn_sid, sciname) %>%
    unique()
  iucn_spp_map <- iucn_spp_cells %>%
    filter(iucn_sid == spp_id$iucn_sid) %>%
    dplyr::select(-parent_sid, -subpop_sid, loiczid = iucn_loiczid) %>%
    group_by(loiczid, iucn_sid) %>%        
    summarize(iucn_area = max(iucn_area))
  ### The group_by() and summarize() are to collapse cells with overlapped polygons, fixing the duplicate 'by' issue
  am_spp_map   <- am_spp_cells %>%
    filter(am_sid == spp_id$am_sid) %>%
    rename(loiczid = am_loiczid)
  spp_map <- full_join(iucn_spp_map, am_spp_map, by = 'loiczid') %>%
    as.data.frame()
  
  return(spp_map)
}
```
***  



```{r,echo=F,eval=FALSE,warning=FALSE,message=F}
spp_list = spp_list%>%
              mutate(am_area = NA,
                     iucn_area=NA,
                     perc = NA,
                     am_perc=NA,
                     iucn_perc=NA)

for (i in 1:nrow(spp_list)){
  
  print(i)
  
  spp <- spp_list[i,]$sciname
  
  spp_map <- get_spp_map(spp)
  
  am_spp_map = spp_map%>%
                dplyr::select(loiczid,am_sid)%>% #add in am_prob when we want to look at probability
                  unique()
  
#   r_am_spp  <-  subs(loiczid_raster, 
#                      am_spp_map[ , c('loiczid', 'am_sid')], 
#                      by = 'loiczid', 
#                      which = 'am_sid', 
#                      subsWithNA = TRUE)
#   
#   spp_list[i,]$am_area = area(r_am_spp,na.rm=T)%>%
#               cellStats(.,stat='sum')
#   
  
  
  iucn_map <- spp_map %>%
    mutate(iucn_pres = iucn_area > 0)%>%
      dplyr::select(loiczid,iucn_area,am_sid,am_prob,iucn_pres)%>%
        unique()
  
  #TODO: incorporate the iucn_area (area of cell that species overlaps with...I think thats what it is)
  
   if(TRUE %in% duplicated(iucn_map$loiczid))next()
#   
#   r_iucn_spp <- subs(loiczid_raster, 
#                      iucn_map[ , c('loiczid', 'iucn_pres')], 
#                     by = 'loiczid', 
#                      which = 'iucn_pres', 
#                      subsWithNA = TRUE)
#   
#   spp_list[i,]$iucn_area = area(r_iucn_spp,na.rm=T)%>%
#                 cellStats(.,stat='sum')
  
  # get percent overlap of map
  
  #iucn map cells
  cells_iucn<-iucn_map%>%
              filter(iucn_pres==TRUE)%>%
              .$loiczid
  cells_am <- am_spp_map%>%
                filter(!is.na(am_sid))%>%
                .$loiczid
  
  cells_total = unique(spp_map$loiczid)
  cells_overlap = intersect(cells_iucn,cells_am)
  
  #percent overlap
  perc = (length(cells_overlap)/length(cells_total))*100
  spp_list[i,]$perc = perc
  
  am_only = setdiff(cells_am,cells_iucn)
  iucn_only = setdiff(cells_iucn,cells_am)
  
  perc_am = (length(am_only)/length(cells_total))*100
  perc_iucn = (length(iucn_only)/length(cells_total))*100
  
  spp_list[i,]$am_perc = perc_am
  spp_list[i,]$iucn_perc = perc_iucn
  
cat(perc,perc_am,perc_iucn)
  
  }


# for Conus arenatus, the IUCN map gives 4 duplicate cells, but each of the entries has a different iucn_area...which doesn't make sense

#row 985 for Acanthurus leucopareius - AM is giving duplicate loiczids but with different am_prob...which is so strange. Looked
# at the raw data and it seems to be in the raw data... (JA on 8/31/15)

write.csv(spp_list,file='spp_list_w_area_percent.csv')
```


###(1) Number of species in both datasets split up by group

```{r,echo=F,warning=F,message=F,warning=FALSE,message=F}
# look at number of species per group

spp_list = read.csv('spp_list_w_area_percent.csv')

df = spp_list%>%
      mutate(diff = iucn_area - am_area)

df <- within(df,spp_group <- factor(spp_group,levels=names(sort(table(spp_group),increasing=TRUE))))

g<-ggplot(df,aes(x=spp_group))+
  geom_bar(stat='bin')+
  theme(axis.text.x=element_text(angle=45,hjust=1,vjust=1))+ coord_flip()+
  labs(x='Species Group', y='Number of Species',title='Number of species in both AquaMaps and IUCN datasets')

print(g)
```

***  

###(2) Looking at the total difference in area (km2) for each species range map in both datasets

```{r,echo=F,warning=F,message=F,warning=FALSE,message=F}

dot = ggplot(df,aes(iucn_area,am_area,color=category))+ geom_point() + geom_abline()+
      labs(x='Area (km2) from IUCN dataset',y='Area (km2) from AquaMaps dataset',title='Comparing total area of species range maps')
print(dot)
```


There are large differences so to get a better sense, here is the same comparison but with area log transformed.
````{r echo=F,warning=F,message=F,warning=FALSE,message=F}

df = df%>%
      mutate(am_log = log(am_area),
             iucn_log = log(iucn_area))

all_status <- ggplot(df,aes(iucn_log,am_log,color=category)) + 
  geom_point() + geom_abline() +
  labs(x='IUCN Area (log transformed)', y='AquaMaps Area (log transformed)',
        title= 'Comparing area of range maps\n (all area values are km2 log transformed)',fill='IUCN Category')

print(all_status)
```

Data split into species groups and by IUCN category
```{r,echo=F,warning=F,message=F}

groups <- ggplot(df, aes(iucn_log, am_log,color=category)) + geom_point() +facet_wrap(~spp_group) + 
        geom_abline() + labs(x='IUCN Area (log transformed)', y='AquaMaps Area (log transformed)',
                             title='Comparing differences in area across species groups', fill='IUCN Category')
print(groups)
```
***  

###(3) Looking at the overlap between species range maps from IUCN and AquaMaps  

This plot shows the total percent of a species range map that is (1) only covered by IUCN cells, (2) only covered by AquaMaps cells and (3) covered by both IUCN and AquaMaps cells. 

####How this breaks down between AquaMaps and IUCN

Here is an example of a single species, the yellow speckled chromis (*Chromis alpha*)  

```{r,echo=F,warning=FALSE,message=F}

spp = 'Chromis alpha'
map <- get_spp_map(spp)

 am_map <- map%>%
            filter(!is.na(am_prob))
    
    
    r_am_spp  <-  subs(loiczid_raster, 
                       am_map[ , c('loiczid', 'am_prob')], 
                       by = 'loiczid', 
                       which = 'am_prob', 
                       subsWithNA = TRUE)
    
    r_am_spp[!is.na(r_am_spp)]<-1
    
    
    iucn_map <- map %>%
                  mutate(iucn_pres = iucn_area > 0)
    
    r_iucn_spp <- subs(loiczid_raster, 
                       iucn_map[ , c('loiczid', 'iucn_pres')], 
                       by = 'loiczid', 
                       which = 'iucn_pres', 
                       subsWithNA = TRUE)
    

#plot
  plot(r_iucn_spp,col='mediumpurple',useRaster=FALSE,axes=FALSE,box=FALSE,legend=FALSE)
  plot(r_am_spp,col='orange',main="Yellow-speckled Chromis", useRaster=FALSE,add=T,axes=FALSE,box=FALSE,legend=FALSE)
  map('world', col = 'gray95', fill = T, border = 'gray80', add = TRUE)
  legend("topright",legend=c("AquaMaps","IUCN"),fill=c('orange','mediumpurple'))
  
```
  

```{r,echo=F,warning=FALSE,message=F}
  
  df <- gather(spp_list,perc.type,percent, perc,am_perc,iucn_perc)%>%
          dplyr::select(sciname,spp_group,perc.type,percent)%>%
            unique()
  
  perc_c = df%>%
            filter(sciname==spp)
  
  p1 = perc_c%>%dplyr::select(percent)
  paste0(round(p1[1,],2),'% of the total area in this map is covered by both IUCN and AquaMaps')
  paste0(round(p1[2,],2),'% of the total area in this map is covered just by AquaMaps')
  paste0(round(p1[3,],2),'% of the total area in this map is covered just by IUCN')
```
  
  
####Now looking across all species, here is the breakdown of how much overlap there is between the two datasets  
  
This plot shows that the majority of species range maps have only ~30% of their full map covered by both maps, and almost no species have more than 50% of their two ranges (IUCN and AquaMaps) overlapping.  

```{r,echo=F,warning=FALSE,message=F}


#look at density plot

all<-ggplot(df,aes(percent,fill=perc.type))+geom_histogram(binwidth=1,alpha=.3)+scale_fill_discrete(name="Legend",labels=c("Overlap", "AquaMaps only", "IUCN only"))+labs(x='Percent of total range', y='Number of Species')

print(all)
```



Split up by species - not sure if this is useful or not
```{r,echo=F}

sp <- ggplot(df,aes(percent,fill=perc.type))+geom_histogram(binwidth=1,alpha=.3) +facet_wrap(~spp_group) +
  scale_fill_discrete(name="Legend",labels=c("Overlap", "AquaMaps only", "IUCN only"))+labs(x='Percent of total range', y='Number of Species')

print(sp)


```
