---
output: html_document
---

``` {r calc species numbers, echo = FALSE}
n_spp_all    <- nrow(spp_all)
n_spp_am     <- nrow(spp_am)
n_spp_am_rev <- nrow(spp_am %>% filter(reviewed == 1))
n_spp_iucn   <- nrow(spp_iucn_maps)
n_spp_both   <- nrow(spp_map_pairs)

p_spp_am     <- round(nrow(spp_am)/n_spp_all * 100, 1)
p_spp_am_rev <- round(n_spp_am_rev/n_spp_am * 100, 1)
p_both_am    <- round(n_spp_both/n_spp_am   * 100, 1)
p_both_iucn  <- round(n_spp_both/n_spp_iucn * 100, 1)
p_both_all   <- round(n_spp_both/n_spp_all  * 100, 1)
```

# Introduction

_Peer beneath the waves anywhere in the world - a moody kelp forest, a bustling coral reefscape, the frigid depths beneath a polar ice sheet, the endless blue of the open ocean - and you will find a unique ecosystem defined as much by which species are absent as by which species are abundant._ Knowing where individuals of a species thrive is fundamental to the sciences of ecology, biogeography, and conservation, among many others.  This knowledge provides foundational information for 
 understanding species ranges and diversity,
 predicting species responses to human impacts and climate change,
 and managing and protecting species effectively.
A rich literature tackles the many dimensions of these questions.

One major outcome of this body of science is the various compiled databases of species distribution maps. Two global-scale repositories predict marine species ranges throughout the world's oceans – AquaMaps modeled species distribution maps (Kaschner et al. 2013) and International Union for Conservation of Nature (IUCN) species range maps (REF).  These two spatial datasets are used for a wide range of purposes, including assessing marine species status (Halpern et al. 2012, Selig et al. 2013), evaluating global biodiversity patterns (Coll et al. 2010, Martin et al. 2014), predicting range shifts (Molinos et al. 2015), and setting conservation priorities (Klein et al. 2015). 

The two datasets ostensibly describe the same information, but significant differences in methodology and intent could lead to dramatically different understandings of our marine ecosystems, with significant implications for policy and conservation recommendations.  Importantly, biases in taxonomic or spatial coverage within a dataset could shift management and conservation actions away from places or species that are most in need.  Inaccurate indications of presence or absence could lead to ineffective marine reserve systems and management plans (Rondinini et al. 2006) (Jetz 2008?).

To understand the implications of differences between the AquaMaps and IUCN datasets, we compared how each data source represents the global spatial and taxonomic distribution of species.  Most notably, AquaMaps includes range maps for many more species (currently `r format(n_spp_am, big.mark = ',')`; `r p_spp_am`% of total), such that most global analyses related to biodiversity to date have used AquaMaps (IUCN range map data exist for only `r format(n_spp_iucn, big.mark = ',')` species).  For the `r format(n_spp_both, big.mark = ',')` (`r p_both_all`% of total) species mapped in both datasets, we examined how well the maps align and determined several issues that lead to misalignment between predicted species distributions, and outline possible improvements. 

We then subjected two recent marine biodiversity studies - an assessment of the status of global biodiversity within the Ocean Health Index (Halpern et al. 2012, 2015) and a global analysis of gaps in protection afforded by marine protected areas (MPAs) (Klein et al. 2015) - to a sensitivity analysis. ... ___substituting___ one dataset over the other, to highlight the possible consequences of different data use decisions on our understanding of _the status of_ marine biodiversity. ___BSH suggestion: We used all unique species range data from both sources but substituted one dataset over the other for shared species to highlight the possible consequences of different data use decisions on our understanding of the status of and protection of marine biodiversity.___

-----
<!--
## Overview of AquaMaps and IUCN datasets  

___BSH comment: Fold this section into the methods and/or Supplementary Methods, or maybe into the results section just below on differences between the datasets – but make the description VERY brief.___

The IUCN publishes species range maps as polygon shapefiles, bundled by taxonomic groups (REF).  Using a geographic information system (GIS), experts outline spatial polygons to represent a given species' extent of occurrence, based on observation records and refined by expert understanding of the species' range and habitat preferences.  _IUCN releases a range map bundle for a taxonomic group once it has been "comprehensively assessed," i.e. at least 90% of the species within the taxonomic group have been evaluated (REF)._  While this mitigates sampling bias within taxa, it also means that entire taxonomic groups remain unavailable until they have met this threshold of comprehensive assessment.  

AquaMaps develops species distribution maps based on modeled relative environmental suitability.  For each mapped species, environmental preferences (e.g. temperature, depth, salinity) are deduced from occurrence records, published species databases such as FishBase, and expert knowledge.  The AquaMaps model overlays these environmental preferences atop a map of environmental attributes on a 0.5 degree grid, creating a global raster of probability of occurrence for each species.  A subset of the resulting distribution maps (`r format(n_spp_am_rev, big.mark = ',')` of `r format(n_spp_am, big.mark = ',')` species maps; `r p_spp_am_rev`%) have been checked or formally reviewed by experts, to validate and improve the parameters and underlying model.

IUCN range maps are intended to include all possible regions in which a species may be present, with the caveat that this does not imply the species is evenly distributed everywhere within the boundaries of the map(REF).  AquaMaps modeled distribution maps provide a more nuanced prediction of species presence within the extent of occurrence, but the model may not capture all the complexities that drive species distribution (REF).  Due to these differences in methodology and intent, IUCN range maps are more likely to overpredict species presence (and underpredict absence) while AquaMaps modeled distribution maps are relatively more likely to underpredict species presence.(REF and REF)
-->

# Results and Discussion

## How and why the datasets differ

The IUCN publishes species range maps as polygon shapefiles, bundled by taxonomic groups (REF).  Using a geographic information system (GIS), experts outline spatial polygons to represent a given species' "limits of distribution" (IUCN 2014) - essentially a refined extent of occurrence, based on observation records and refined by expert understanding of the species' range and habitat preferences.  AquaMaps models species distribution maps based environmental preferences (e.g. temperature, depth, salinity) deduced from occurrence records, published species databases such as FishBase, and expert knowledge.  The AquaMaps model overlays these environmental preferences atop a map of environmental attributes on a 0.5 degree grid, creating a global raster of probability of occurrence for each species.

The two datasets have notably different taxonomic (Fig. 1A) and regional (Fig. 1B,C) coverage. IUCN-mapped species focus on tropical latitudes and away from the Atlantic and Eastern Pacific compared to AquaMaps-mapped species. These differences can be mitigated by using both datasets, but the underlying methodological differences complicate such direct comparisons, and furthermore ___requires resolving differences (or choosing one source over the other) for the `r format(n_spp_both, big.mark = ',')` shared species.___

-----


``` {r load_spp_list_by_quad, echo = FALSE}
spp_list_quads <- read.csv('data/spp_list_quads.csv', stringsAsFactors = FALSE)
n_q1 <- nrow(spp_list_quads %>% filter(quad == 'q1'))
n_q2 <- nrow(spp_list_quads %>% filter(quad == 'q2'))
n_q2_am_sm  <- nrow(spp_list_quads %>% filter(sm_range == 'AM' & quad == 'q2'))
n_q2_corals <- nrow(spp_list_quads %>% filter(str_detect(tolower(spp_group), 'coral') & quad == 'q2'))
n_corals    <- nrow(spp_list_quads %>% filter(str_detect(tolower(spp_group), 'coral')))
n_q3 <- nrow(spp_list_quads %>% filter(quad == 'q3'))
n_q4 <- nrow(spp_list_quads %>% filter(quad == 'q4'))
```

___BSH suggestion___ To explore differences in species distribution and range between the two datasets, we plotted the distribution alignment (where on the map) against the ratio of extent (how much of the map) for each shared species (Fig. 2a). __Overall we see a negative linear relationship, where increasing similarity in extent (extent ratio approaches 100%) correlates with a decrease in alignment.  [explain/discuss].__  

Dividing the map-paired species into quadrants highlights different categories of relationships. The upper right quadrant includes species (n = `r n_q1`) whose map pairs agree in both spatial distribution and extent of described ranges. These species tend to be well-studied and predominantly fall within wide-ranging pelagic organisms such as marine mammals, tunas, and billfishes (Fig. 2B). This result is not surprising, as species with very large ranges are likely to be more aligned regardless of methodology simply because they exist nearly everywhere.

The extent-mismatched maps contained in the upper left quadrant (n = `r n_q2`) provide a richer set of examples to understand the fundamental differences between the datasets.  For many of these species, while the spatial distribution is similar, the IUCN range predicts a significant buffer zone beyond the AquaMaps range.  Corals dominate this quadrant (`r n_q2_corals`; `r round(n_q2_corals/n_q2 * 100, 1)`% of all species in this quadrant).  Most corals and reef-associated organisms prefer shallower waters, and we suspect that many of these disagreements in extent can be explained by bathymetry: seafloor depth is explicitly included in AquaMaps models, while depth is frequently ignored as a factor in IUCN range maps, leading to overprediction of species range into areas too deep to support the species. (For examples and details, see SOM) 

The differences in extent of species maps found in the lower right quadrant (n = `r n_q3`) can in many cases be attributed to "two wrongs make a right."  For these species, IUCN ranges frequently overextend into unsuitable depths, as in the case of many upper left quadrant species, while at the same time AquaMaps ranges often aggressively extrapolate presence into locations where IUCN predicts absence.  This results in an area ratio close to 100%, but similar extents are meaningless when the distributions are poorly aligned.

Moving from the lower right to lower left quadrant, AquaMaps more frequently extrapolates species presence well beyond known occurrences, resulting in poor matches in both distribution and extent.  Of the `r n_q4` species in this quadrant, 159 (39.7%) are data-poor (in which the AquaMaps environmental envelope model is based on fewer than 10 known cell occurrences), compared with 21.1% for the full set of map-paired species and only 8.2% of species in the upper left quadrant.  A mere 6.2% of AquaMaps maps in this quadrant have undergone expert review, compared with 13.4% of all map-paired species and 24.9% of species in the upper left quadrant.

This breakdown by quadrant helps explain the negative linear relationship seen in the plot, in which increasing similarity in extent correlates with decreasing distribution alignment.  Expanding species range beyond known occurrences, the marginal range predicted by the AquaMaps model will fall in different locations than the marginal range predicted using IUCN methodology.  For species with dissimilar extents, predicted distribution for the smaller range can more easily fall within the generous bounds of the larger range.  As we examine species with increasingly similar extents, the differences in methodology become more difficult to "hide," and the distribution alignment generally becomes poorer.

-----

# Implications of the differences

___From JSL:___ _I think you need a broader discussion of implications. How many times has IUCN/Aquamaps been cited? How long have they been around? What scales are the studies? Then you can muse about how results/interpretations would be different depending on which dataset they used, and reference Fig 3’s._

_Then, say for example, the OHI uses these datasets together (don’t need to get too much into OHI background I think, maybe they don’t even need to know about the biodiversity goal?), but say that to try to combine them they use a threshold at 40% since others use this threshold (but check out the SOM for analyses about thresholds)_

## Application to OHI

The global Ocean Health Index (OHI) (Halpern et al. 2012, 2015), a composite index made up of 10 goals, uses species spatial distribution data and IUCN Red List conservation status to calculate biodiversity status for each of 221 exclusive economic zones. To maximize the number of represented species, spatial distributions were gleaned from both IUCN and AquaMaps datasets, preferring IUCN data for species in both sources.  OHI uses a probability threshold of 40% to determine species presence for AquaMaps data.  Species with no spatial data in either dataset were excluded, as were species with insufficient information to determine conservation status (including species listed as not evaluated or data deficient).

_Briefly summarize results_

Since the Ocean Health Index Species subgoal relies on spatial data from both datasets, the impacts of these threshold and preference changes will be somewhat muted (<a href = '#fig3'>Figure 3</a>).  When IUCN range maps are the preferred data source, only the subset of AquaMaps-only species will be affected by threshold changes; and when AquaMaps is the preferred source, the IUCN-only species will dampen the effect of a threshold change.

-----  

## Application to MPA Gap Analysis

Klein et al. (2015) compared the global distribution of species to the global distribution of marine protected areas to assess how well the MPAs protect key species, and to identify which species fall through the gaps in protection.  The MPA gap analysis relied on the AquaMaps database available in early 2015, comprising distribution maps for XXX species.  Species presence was defined using a probability threshold of 50% or greater.  The study defined four subsets of protected areas using the World Database of Protected Areas, based upon combinations of IUCN protected area classification, marine designation, and spatial marine coverage.  The study determined that currently defined marine protected areas (IUCN categories I through IV overlapping with marine areas) are woefully inadequate in protecting marine biodiversity; 90.5% of species have less than 5% of their overall range represented within marine protected areas.

working on the analysis...

* rerun based on old AquaMaps at 50% (to verify methodology gets a close result to paper), then new AquaMaps for updated baseline
    * consider also: filter on STATUS == 'Designated', and filter out those pesky fishery management areas and manatee speed zones...
* run with threshold at 0% (not all the same as original paper)
* run with filtering out data-poor species?
* run with IUCN maps instead of AquaMaps
* Which aspects of the analysis should we focus on?:
    * overall protection, for 4th subset of protected areas? or overall for all subsets? (see paper, table 1)
    * overall vs broken down by taxonomic groups? (see paper, figure 1)
    * spatial distribution of gap species (see paper, figure 2)

_Predictions:  What happens when using a 0% threshold instead of a 50%?  What happens when using IUCN rather than AM?  IUCN tends to overestimate extent esp for coastal species?  More species will be included in MPAs so fewer apparent gap species.  Included range area inside MPAs will increase.  Policy implications?_

-----

# Conclusions: What can be done to address the differences

_While the differences between species distributions described by AquaMaps and IUCN can drive different conclusions for conservation research, some of the mechanistic causes can be addressed._ ___oh god this section is so passive voice... arrrrgghhhh___

__For data providers:__

* The training manual for experts creating IUCN Red List maps for marine species suggests that "bathymetry can be used to delineate species' ranges limited by depth in the same way as elevation is used for terrestrial species" (REF https://www.conservationtraining.org/mod/scorm/player.php?a=1412&currentorg=M5L1_ORG&scoid=6492).  Unfortunately, this recommendation appears to be applied haphazardly.  Especially for demersal communities dependent on photosynthesis, bathymetry should be a primary consideration; even a low resolution cutoff of 200 meters would improve the utility of species range maps.

* The AquaMaps model is a powerful tool for quickly estimating species presence based on limited information; however, the model's output is only as good as the input data, which for marine species is often sorely limited.  AquaMaps encourages experts to review and comment on the predicted species distribution maps; lessons from the small sample of reviewed maps can be generalized, for example to other species in the same taxonomic group, to further refine the predictions, especially for data-poor species.

* In addition to environmental preferences and conditions, AquaMaps mapping parameters include area restrictions to help limit over-extrapolation of AquaMaps models.  Currently the area restrictions are denoted by FAO Major Fishing Areas (REF find a reference in AquaMaps somewhere); however, allowing area restrictions based on biogeographical criteria, a la MEOW, would likely have better resolution and predictive power, especially for data-poor species.

__For data users:__

* understand that range maps a la IUCN tend to overestimate presence; AquaMaps are less likely to overestimate presence, but trade off for overestimating absence. Both better than simple point locality data. (Rondinini etc)
* Using both datasets together can increase the taxonomic and spatial breadth of coverage, as long as the differences between the datasets can be reasonably minimized.  Consider:
    * An AquaMaps presence threshold of 0% more closely approximates the "limits of distribution" outlined by IUCN range maps.
    * Additional filters and criteria on AquaMaps distribution maps can help avoid overextrapolation for data poor species.  The AquaMaps native range dataset includes information on expert review and the number of occurrence cells used to generate the environmental envelope.
    * For depth-limited species or taxa, clipping IUCN range polygons to a reasonable bathymetry contour can limit overprediction of species presence into unsuitable habitats.

