---
title: 'OHI: Species: ingest IUCN shapefiles'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ohiprep/src/templates/ohi_hdr.html'
  pdf_document:
    toc: true
---

``` {r setup, echo = TRUE, message = FALSE, warning = FALSE}

knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'Figs/',
                      echo = TRUE, message = FALSE, warning = FALSE)

library(rgdal)
library(maptools)
library(raster)
library(tidyverse)
library(stringr)

dir_M <- c('Windows' = '//mazu.nceas.ucsb.edu/data_edit',
           'Darwin'  = '/Volumes/ohi',
           'Linux'   = '/home/shares/ohi')[[ Sys.info()[['sysname']] ]]

dir_git <- '~/github/IUCN-AquaMaps'
  
dir_anx <- file.path(dir_M, 'git-annex/globalprep/spp_ico/vAM_IUCN')
dir_data_remote <- file.path(dir_M, 'git-annex/globalprep/_raw_data')

source(file.path(dir_git, 'data_explore/data_explore_fxns.R'))

```

# Methods

## Ingest shapefiles from IUCN into half-degree cells

``` {r define_paths}

iucn_shp_dir <- file.path(dir_data_remote, 'iucn_spp', 'd2015', 'iucn_shp')

cache_dir <- (file.path(dir_anx, 'iucn_intersections'))

rast_file <- file.path(dir_anx, '../rgns/loiczid_raster.tif')
```

``` {r define_functions}

extract_from_shp <- function(shp, cache_dir, rast, 
                             fn_tag   = NULL,
                             reload   = FALSE) {
  ### will extract a shapefile to a raster, and save the .csv as a
  ### file in cache_dir with a matching name (e.g. BOTW.shp -> BOTW.csv).
  ### NOTE: any filtering of shapefiles should be done before passing them
  ### to this function; e.g. filter TERRESTRIAL_MAMMALS.shp to just those
  ### that occur in marine habitats.
  ### * shp must include .shp extension and must include full path name.
  
  if(!file.exists(shp)) {
    message('Shapefile ', shp, ' does not exist.  shp argument must contain ',
            'the full path name, including .shp extension.')
    
    return(data.frame(msg = 'pathname error'))
  }

  ### determine species group (the shapefile base name without extension)
  spp_gp <- basename(shp) %>% str_replace('.shp', '')
  
  cache_file <- file.path(cache_dir, 
                          paste0(spp_gp, 
                                 ifelse(is.null(fn_tag), '', fn_tag),
                                 '.csv'))
  
  ### if reload == FALSE, and the file exists, don't waste your friggin' time here, move on to next group.
  if(file.exists(cache_file) & reload == FALSE) {
    message(sprintf('IUCN <-< LOICZID lookup file already exists for species group %s;',
                    ' file location:\n  %s', spp_gp, cache_file))
    return(data.frame(msg = paste(cache_file, ' already exists, not reprocessed')))
  } else {
    ptm <- proc.time()
    fsize <- round(file.size(shp)/1e6, 2)
    message(sprintf('Reading species group shapefile %s, %.2f MB\n  %s', spp_gp, fsize, shp))

    ### Because the IUCN metadata states that shapefiles are unprojected lat-long with WGS84,
    ### use readShapePoly (rather than readOGR) and manually tell it the projection...
    spp_shp <- maptools::readShapePoly(fn = shp,
                                       proj4string = CRS('+init=epsg:4326'),
                                       delete_null_obj = TRUE)
    # spp_shp <- rgdal::readOGR(dsn = dirname(shp),
    #                           layer = basename(shp) %>% str_replace('.shp', ''),
    #                           stringsAsFactors = FALSE)
    message(sprintf('Elapsed read time: %.2f seconds', (proc.time() - ptm)[3]))

    ### standardize some variable names; this is mostly for BOTW
    names(spp_shp@data) <- tolower(names(spp_shp@data))
    names(spp_shp@data)[names(spp_shp@data) %in% c('id_no', 'sisid')] <- 'iucn_sid'
    names(spp_shp@data)[names(spp_shp@data) %in% c('binomial')]       <- 'sciname'
    names(spp_shp@data)[names(spp_shp@data) %in% c('presenc')]        <- 'presence'
    ### corals 1 (and others?) doesn't have subpop variable; add NAs
    if(!'subpop' %in% names(spp_shp@data))
      spp_shp@data$subpop <- NA


    message(sprintf('... processing %s species',
                    length(unique(spp_shp@data$iucn_sid))))

    message('Extracting polygons to LOICZID cells...')
    ptm <- proc.time()
    spp_shp_prop <- raster::extract(rast, spp_shp, 
                                    weights = TRUE, normalizeWeights = FALSE, 
                                    progress = 'text')
    message(sprintf('Elapsed process time: %.2f minutes', (proc.time() - ptm)[3]/60))

    
    ### combine sciname, iucn_sid, presence, and subpop code for a single unique identifier
    shp_id <- data.frame('sciname'  = spp_shp@data$sciname,
                         'iucn_sid' = spp_shp@data$iucn_sid,
                         'presence' = spp_shp@data$presence,
                         'subpop'   = spp_shp@data$subpop) %>%
      unite(name_id, sciname, iucn_sid, presence, subpop, sep = '_')

    names(spp_shp_prop) <- shp_id$name_id

    ### convert list to data frame.
    spp_shp_prop_df <- plyr::ldply(spp_shp_prop, rbind)
    spp_shp_prop_df <- spp_shp_prop_df %>%
      rename(name_id   = .id,
             LOICZID   = value,
             prop_area = weight) %>%
      separate(name_id, c('sciname', 'iucn_sid', 'presence', 'subpop'), sep = '_') %>%
      distinct()

    ### save .csv for this species group
    message(sprintf('%s: %s species maps, %s total cells in output file',
                    spp_gp, length(unique(spp_shp_prop_df$iucn_sid)),
                    nrow(spp_shp_prop_df)))
    message(sprintf('Writing IUCN<->LOICZID intersection file for %s to:\n  %s', spp_gp, cache_file))
    write_csv(spp_shp_prop_df, cache_file)
  }
  
  return(invisible(spp_shp_prop_df))
}

```


### Extract IUCN shapefiles to csvs.

All shapefiles within the IUCN shapefile folder are extracted against the LOICZID raster from AquaMaps.  The resulting dataframes are saved as a CSV for each species group (each shapefile).

NOTE: reload is disabled.  To re-extract a shapefile, simply delete the existing .csv extracted from .shp

### available shape files:

``` {r display_shps, results = 'asis'}

shp_files <- c(list.files(iucn_shp_dir,
                        pattern = '.shp$',
                        full.names = TRUE))

csv_files <- list.files(file.path(dir_anx, 'iucn_intersections'),
                        pattern = '.csv$') %>%
  str_replace('.csv$', '')

shp_files_to_do <- shp_files[!(basename(shp_files) %>% str_replace('.shp$', '') %in% csv_files)]

shp_info <- lapply(shp_files, file.info) %>%
  bind_rows() %>%
  mutate(shp_file = shp_files %>% str_replace('/home/shares/ohi/git-annex/globalprep/', ''),
         size_mb  = round(size / 1e6, 2)) %>%
  dplyr::select(shp_file, size_mb, mod_time = mtime, user = uname, group = grname) %>%
  mutate(to_process = basename(shp_file) %in% basename(shp_files_to_do))

DT::datatable(shp_info)

```


``` {r extract_shps_to_csv_series}

spp_list <- read_csv(file.path(dir_git, 'data/spp_iucn_maps_verified.csv'))

rast <- raster(rast_file)

reload <- FALSE

shps_done <- list.files(cache_dir) %>%
  str_replace('.csv', '.shp') %>%
  paste(collapse = '|') ### converts multiple strings into single regex pattern

if(shps_done == '') {
  shps_remain <- shp_files
} else {
  shps_remain <- shp_files[!str_detect(shp_files, shps_done)]
}

for (shp in shps_remain) {
  # shp <- shps_remain[1]
  shp_csv <- extract_from_shp (shp,
                               cache_dir = cache_dir,
                               rast = rast,
                               reload = reload)
}

```


``` {r check_that_all_files_processed}
### run this bit to check whether all files processed properly.  Errors
### seem to typically knock a few out of the running for some reason.
shps_done <- list.files(cache_dir) %>%
  str_replace('.csv', '.shp') %>%
  paste(collapse = '|') 
    ### converts multiple strings into single regex pattern so str_detect can
    ### test for patternA OR patternB OR patternC etc at one swoop
shps_remain <- shp_files[!str_detect(shp_files, shps_done)]

if (length(shps_remain > 0)) {
  message('Shapefiles not processed: \n  ', paste(shps_remain, collapse = '\n  '))
  message('... Reprocessing these files!')
  shp_df_list <- mclapply(shps_remain, 
                            FUN = extract_from_shp,
                                    cache_dir = cache_dir,
                                    rast = rast,
                                    reload = reload,
                          mc.cores = 12) 
  
} else {
  message('All shapefiles processed! yay!')
}
```

### available CSV files:

``` {r display_csvs, results = 'asis'}

csv_files <- list.files(cache_dir, pattern = '.csv$', full.names = TRUE)

cat('All .csv files located in ', cache_dir, '\n')

csv_info <- lapply(csv_files, file.info) %>%
  bind_rows() %>%
  mutate(basename = basename(csv_files),
         size_mb  = round(size / 1e6, 2)) %>%
  dplyr::select(basename, size_mb, mod_time = mtime, user = uname, group = grname)

DT::datatable(csv_info)

```

## Combine all extracted CSVs

``` {r combine_all_extracted_csvs}

spp_list <- read_csv(file.path(dir_git, 'data/spp_iucn_maps_verified.csv'))

csv_list <- list.files(cache_dir, pattern = '.csv$', full.names = TRUE)

spp_cells_df <- lapply(csv_list, read_csv) %>%
  bind_rows()

names(spp_cells_df) <- tolower(names(spp_cells_df))

spp_missing <- spp_list %>%
  filter(!iucn_sid %in% spp_cells_df$iucn_sid) %>%
  distinct()
### none missing!

spp_extra <- spp_cells_df %>%
  filter(!iucn_sid %in% spp_list$iucn_sid) %>%
  dplyr::select(-loiczid, -prop_area, -presence) %>%
  distinct()
### no extras!

spp_cells_file <- file.path(dir_anx, 'int', 'iucn_spp_cells_d2015.csv')
write_csv(spp_cells_df, spp_cells_file)

spp_cell_file_info <- file.info(spp_cells_file) %>% 
  mutate(size_mb = size / 1e6 %>% round(2)) %>%
  dplyr::select(size_mb, mod_time = mtime, user = uname, group = grname)

```

File saved at: `r spp_cells_file`

`r knitr::kable(spp_cell_file_info)`

Extra species?  `r nrow(spp_extra)`

Missing species?  `r nrow(spp_missing)`

-----

